---
format: gfm
---

## Which nonparametric test for the median is best?

There are two nonparametric tests for the median. Which test is best? Is one test better in some situations and the other test better in another situation? This repo compares the two hypothesis tests with data from four different distributions.

Tests:

-   Empirical Likelihood
-   Sign Test

Distributions:

-   Gaussian
-   Cauchy
-   Uniform
-   Exponential

The main outcome is type I error is the basically the same for both tests. Type II error is slightly lower for the empirical likelihood test.

## Experimental Design

In this experiment, two settings vary. N goes from 25 to 200 increasing by 25. The distribution changes between normal, cauchy, uniform and exponential. Each combination is repeated 100 times. Per row, the two test receive the exact same data and have the exact same null hypothesis.

The first few simulations look like 
```{r}
#| include: false
library(tidyverse)
library(scales)
```

```{r}
#| echo: false
typeI <- readRDS("data/typeI.rds") %>%
  select(dist, n, iteration, LR_P, S_P) %>%
  mutate(
    LR_P = round(LR_P, 4),
    S_P = round(S_P, 4)
  ) %>%
  arrange(dist, n, iteration)
typeI %>%
  print(n = 10)
```


## Type I Error

Each combination of N and distribution is aggregated to calculate type I error. 

For all four distributions, type I error is essentially the same.

```{r}
#| echo: false
#| warning: false
#| message: false
calc_type_I <- function(pvalues) {
  out <- mean(pvalues <= .05)
}

typeI <- readRDS("data/typeI.rds") %>%
  select(dist, n, iteration, LR_P, S_P) %>%
  mutate(
    LR_P = round(LR_P, 4),
    S_P = round(S_P, 4)
  ) %>%
  arrange(dist, n, iteration)

typeI_02 <- typeI %>%
  group_by(dist, n) %>%
  summarise(across(contains("_P"), calc_type_I)) %>%
  ungroup() %>%
  pivot_longer(cols = contains("_P"), values_to = "typeI", names_to = "test") %>%
  mutate(
    typeI = round(typeI, 2),
    test = if_else(test == "LR_P", "Likelihood", "Sign Test")
  )

graph <- ggplot(data = typeI_02, mapping = aes(x = n, y = typeI, color = test)) +
  geom_point() +
  scale_x_continuous(breaks = seq(25, 200, 25)) +
  scale_y_continuous(labels = percent, breaks = seq(0, 1, .2), limit = c(0, 1)) +
  geom_hline(yintercept = .05) +
  labs(x = "N", y = "Type I") +
  facet_wrap(~dist)

graph
```
## Type II Error
For type II error, the simulation is repeated. The major change is the null hypothesis is false. Each combination of N and distribution is aggregated to calculate type II error. Note the effect size is only .25.

Type II error is slightly lower for the empirical test for all four distributions.

```{r}
#| echo: false
#| warning: false
#| message: false
calc_type_II <- function(pvalues) {
  out <- mean(pvalues > .05)
}

typeII <- readRDS("data/typeII.rds") %>%
  select(dist, n, iteration, LR_P, S_P) %>%
  mutate(
    LR_P = round(LR_P, 4),
    S_P = round(S_P, 4)
  ) %>%
  arrange(dist, n, iteration)

typeII_02 <- typeII %>%
  group_by(dist, n) %>%
  summarise(across(contains("_P"), calc_type_II)) %>%
  ungroup() %>%
  pivot_longer(cols = contains("_P"), values_to = "typeII", names_to = "test") %>%
  mutate(
    typeII = round(typeII, 2),
    test = if_else(test == "LR_P", "Likelihood", "Sign Test")
  )

graph <- ggplot(data = typeII_02, mapping = aes(x = n, y = typeII, color = test)) +
  geom_point() +
  scale_x_continuous(breaks = seq(25, 200, 25)) +
  scale_y_continuous(labels = percent, breaks = seq(0, 1, .2), limit = c(0, 1)) +
  labs(x = "N", y = "Type II") +
  facet_wrap(~dist)

graph
```
